### 2022-09-14 hw3-1/anshit/HW_3.Rmd

@@

```{r}
data = read.table('./temps.txt', sep="\t", dec=".", header = TRUE)
data
```

@@

>>>>

```{r}
library(grid)
```

@@

```{r}
data = read.table('../temps.txt', sep="\t", dec=".", header = TRUE)
data
```

@@

```{r}
### Transform to celsius
temp_cols = names(data)[!(names(data) %in% c("DAY"))]
#Transform to celsius
data_celsius = data
data_celsius[temp_cols] <- lapply(data_celsius[temp_cols], function(f) (f-32)*5/9)
```

```{r}
CUSUM_ad = function(x, year, days = data$DAY ,C = 0.30, T = 4, if_c_relative = FALSE){
  mean_x = mean(x)
  sd_x = sd(x)
  if (if_c_relative) {
    C = C * sd_x
  }
  #apply cusum
  x1 = lapply(x, function(xi)  (mean_x - xi - C))
  st_df = data.frame(st=double())
  st = 0
  for(i in 1:length(x)){
    #print (paste(i,class(st),class(x1[[i]])))
    st = max(0, st + x1[[i]])
    st_df[i,] = c(st)
  }
  #Find index where St > T
  st_df$Day = days
  st_df["anamoly"] = 0
  st_df$anamoly[st_df$st > T] = 1
  st_df$xt = x
  #plot x with T and if decrease see
  par(mfrow=c(2,1))
  plot(st_df$x)
  title(paste("Temperature (in Celsius) and S_t (bottom) for year",year), line = -1, outer = TRUE)
  plot(st_df$st)
  pushViewport(viewport())
  grid.lines(x = c(0,1), y = grconvertY(T, "user", "ndc"), gp = gpar(col = "red"))
  change_detected = which.max(st_df$st>T)
  grid.lines(x = grconvertX(change_detected, "user", "ndc"), y = c(0,1), gp = gpar(col = "red"))
  popViewport()
  #return first point where St touches T
  change_length = length(days)-change_detected
  return (c(as.numeric(year),as.numeric(change_detected),as.numeric(change_length)))
}
```

```{r}
### Run Cusum over data_celsius for all the years
winter_start_days = lapply(temp_cols, function(col) 
  CUSUM_ad(x=data_celsius[,col], year = substr(col,2,5), C = 4, T = 25)
  )
```

```{r}
change_df <- as.data.frame(do.call(rbind, winter_start_days))
colnames(change_df) = c("year","winter_start","winter_length")
change_df["year_bucket"] = "a.1996_2000"
change_df[change_df$year>=2000,"year_bucket"] = "b.2000_2004"
change_df[change_df$year>=2004,"year_bucket"] = "c.2004_2008"
change_df[change_df$year>=2008,"year_bucket"] = "d.2008_2012"
change_df[change_df$year>=2012,"year_bucket"] = "e.2012_2016"
change_df
```

```{r}
aggregate(cbind(winter_start,winter_length) ~ year_bucket, change_df, mean)
```

```{r}
month_df <- data.frame(data)
month_df["month"] = "month"
month_df[grepl( "Jul", month_df$DAY), "month"] = "Jul"
month_df[grepl( "Aug", month_df$DAY), "month"] = "Aug"
month_df[grepl( "Sep", month_df$DAY), "month"] = "Sep"
month_df[grepl( "Oct", month_df$DAY), "month"] = "Oct"
month_df = month_df[,-1]
for ( col in 1:ncol(month_df)){
    colnames(month_df)[col] <-  sub("X", "", colnames(month_df)[col])
}
month_df
```

```{r}
year_bucketing_function = function(df){
  df["year_bucket"] = "a.1996_2000"
  df[df$Year>=2000,"year_bucket"] = "b.2000_2004"
  df[df$Year>=2004,"year_bucket"] = "c.2004_2008"
  df[df$Year>=2008,"year_bucket"] = "d.2008_2012"
  df[df$Year>=2012,"year_bucket"] = "e.2012_2016"
  df
}
```

```{r}
aggregate_on_month_and_transpose = function(df, func){
  return_df = aggregate(. ~ month, df, func)
  n = return_df$month
  # transpose all but the first column (month)
  return_df <- as.data.frame(t(return_df[,-1]))
  colnames(return_df) <- n
  return_df = cbind(Year = rownames(return_df), return_df)
  rownames(return_df) = 1:nrow(return_df)
  return_df = return_df[, c(1, 3, 2, 5, 4)]
  return_df
}
```

```{r}
mean_df = aggregate_on_month_and_transpose(month_df, mean)
mean_df = year_bucketing_function(mean_df)
mean_df = mean_df[,-1]
mean_df = aggregate(.~year_bucket, mean_df, mean)
mean_df
```

```{r}
min_df = aggregate_on_month_and_transpose(month_df, min)
min_df = year_bucketing_function(min_df)
min_df = min_df[,-1]
min_df = aggregate(.~year_bucket, min_df, mean)
min_df
```

```{r}
max_df = aggregate_on_month_and_transpose(month_df, max)
max_df = year_bucketing_function(max_df)
max_df = max_df[,-1]
max_df = aggregate(.~year_bucket, max_df, mean)
max_df
```
